\pkg{hesim}: Health Economic Simulation Modeling and Decision Analysis

### 2020-12-17 manuscript.Rnw
First draft of iCTSTM example

<<include=FALSE>>=
library(knitr)
knitr::render_sweave()
opts_chunk$set(
  engine="R", tidy = FALSE, eval = TRUE, cache = TRUE,
  fig.height = 5, fig.width = 7
)
@

@@

%% -- Illustrative example -----------------------------------------------------
\section{Illustrative example} \label{sec:illustrative-example}
We consider a three state model commonly used describe disease progression in oncology. As shown in Figure~\ref{fig:three-state-model}, the three health states are stable disease (i.e. not progressed), progressed, and death. Patients can transition to a more severe health state (stable $\rightarrow$ progressed, stable $\rightarrow$ death, and progression $\rightarrow$ death) but cannot recover to a less severe health state.

@@

Multi-state models can be fit in one of two ways. First, a joint survival model with interaction terms for different transition can be estimated, which is useful if there are constraints in the parameters across
transitions, such as coefficients that are assumed equal across transitions. Here, we will take a second approach that is computationally more efficient and fit a separate model for each transition. We illustrate with parametric Weibull models but multiple distributions should be compared in a real application \citep{williams2017cost}. The treatment effect is assumed to only influence transitions from stable disease since patients are assumed to switch to a second line treatment after progression.

@@

@@

@@

<<ictstm_drugcost_setup, echo=FALSE, results="hide">>=
n_times <- 2
n_states <- nrow(states)
n_strategies <- nrow(strategies)
n_i <- n_states * n_times
drugcost_dt <-  data.table(
    strategy_id = rep(strategies$strategy_id, each = n_i),
    state_id = rep(rep(states$state_id, each = n_times), n_strategies),
    time_start = rep(c(0, 3/12), n_states * n_strategies),
    est = c(rep(2000, n_i -1), 1500, 
            rep(12000, n_i -1), 1000,
            rep(15000, n_i - 1), 1200
            ) 
)
@

<<ictstm_drugcost>>=
drugcost_tbl <- stateval_tbl(
  drugcost_dt,
  dist = "fixed",
  hesim_data = hesim_dat
)
@

\subsubsection{Simulation}
<<ictstm_n_samples>>=
n_samples <- 100
@

@@

@@

@@

@@

@@

@@

@@

>>>>

<<include=FALSE>>=
library(knitr)
knitr::render_sweave()
opts_chunk$set(
  engine="R", tidy = FALSE, eval = TRUE, cache = TRUE,
  fig.height = 4, fig.width = 6
)
@

@@

%% -- Illustrative example -----------------------------------------------------
\section{Illustrative example} \label{sec:illustrative-example}
We consider a three state model commonly used describe disease progression in oncology. As shown in Figure~\ref{fig:three-state-model}, the three health states are stable disease (i.e. not progressed), progressed, and death. Patients can transition to a more severe health state (stable $\rightarrow$ progressed, stable $\rightarrow$ death, and progression $\rightarrow$ death) but cannot recover to a less severe health state. We assume that patients remain on first line (1L) treatment until progression, at which time they switch to a second line (2L) treatment.

@@

Multi-state models can be fit in one of two ways. First, a joint survival model with interaction terms for different transition can be estimated, which is useful if there are constraints in the parameters across
transitions, such as coefficients that are assumed equal across transitions. Here, we will take a second approach that is computationally more efficient and fit a separate model for each transition. We illustrate with parametric Weibull models but multiple distributions should be compared in a real application \citep{williams2017cost}. The treatment effect is assumed to only influence transitions from stable disease since patients are assumed to switch to a 2L treatment after progression.

@@

Utility and cost values are stored in \code{stateval_tbl} objects. Utility is assumed to equal 0.8 with stable disease (standard error (SE) = 0.02) and 0.6 (SE = 0.05) with progressed disease. Utility values are assumed to follow a beta distribution, which can be parameterized either using its shape parameters or via the mean and standard error (in which case the method of moments is used to derive the shape parameters).

@@

Both medical and drug costs are considered. Medical costs are assumed to follow a gamma distribution which is often appropriate as they tend to be right skewed. Like utility, the distribution is characterized by the mean (\$2,000 with stable disease, \$9,500 with progressed disease) and standard error (assumed equal to the mean) and the method of moments is used to derive the underlying shape and scale parameters.

@@

Drug costs are assumed to be fixed (i.e., measured without uncertainty). Costs are \$2,000 with SOC, \$12,000 with New 1, and \$15,000 with New 2 when on 1L treatment. All patients are assumed to switch to the same treatment at 2L after progression, which costs \$1,500 for the first 3 months and \$1,200 thereafter.

<<ictstm_drugcost_setup, echo=FALSE, results="hide">>=
n_times <- 2
n_states <- nrow(states)
n_strategies <- nrow(strategies)
n_i <- n_states * n_times
drugcost_dt <-  data.table(
    strategy_id = rep(strategies$strategy_id, each = n_i),
    state_id = rep(rep(states$state_id, each = n_times), n_strategies),
    time_start = rep(c(0, 3/12), n_states * n_strategies),
    est = c(2000, 2000, 1500, 1200,
           12000, 12000, 1500, 1200,
           15000, 15000, 1500, 1200)
) 
@

<<ictstm_drugcost>>=
drugcost_tbl <- stateval_tbl(
  drugcost_dt,
  dist = "fixed",
  hesim_data = hesim_dat
)
print(drugcost_tbl)
@

\subsubsection{Simulation}
Before constructing the economic model, we specify the number of parameters samples that will be drawn for each submodel for the PSA.

<<ictstm_n_samples>>=
n_samples <- 100
@

The input data for the transition model consists of one row for each treatment strategy and patient and can be easily generated from a \code{hesim_data} object with the \code{expand} function.

@@

The transition model is then constructed as a function of the parameters (from the fitted Weibull models) and the input data. We must also specify the allowed transitions, the desired number of PSA samples (drawn from the multivariate normal distribution of the maximum likelihood estimate of the regression coefficients), the "clock" (recall that we used a clock reset approach when fitting the Weibull models), and the starting age of the patients. The starting age of the patients is useful because it allows us to specify a maximum age for patients during the simulation, ensuring that the simulated patients do not live unrealistically long lives.

@@

The utility and cost models can be easily constructed from the \code{stateval_tbl} objects with the \code{create_StateVals()} function, in which the (transformed) parameters are \code{tparams_mean} objects. The \code{hesim_data} object provides the information required to ensures that state values are constant within groups not specified within a particular \code{stateval_tbl}; for instance, since utility only varies by health state, it is assumed constant across patients, treatment strategies, and time. The models for each cost category are combined in a list.

@@

The complete economic model is an constructed by combining the transition, utility, and cost submodels.

@@

Once the economic model has been constructed, it is straightforward to simulate outcomes. A trajectory through the multi-state model is simulated with the \code{$sim_disease()} method with patients assumed to live to a maximum age of 100.

@@

State probabilities (averaged across patients in a given subgroup) can be constructed from the disease progression simulation output with the \code{$sim_stateprobs()} method. As shown in Figure~\ref{fig:ictstm-stprobs}, patients using the newer treatments remain in stable disease longer and have longer expected survival.

@@

QALYs and costs are simulated from the disease progression simulation output using the utility and cost submodels with the \code{$sim_qalys()} and \code{$sim_costs()} methods. Outcomes are simulated for each discount rate specified with the \code{dr} argument. By default, \code{$sim_qalys()} will return life-years (i.e., a utility value of 1 for each year of life) in addition to QALYs.

@@

For a given subgroup, QALYs and costs are simulated for each PSA sample, treatment strategy, and health state. They are summarized (i.e., averaged) across PSA samples and treatment strategies (and subgroups) with the \code{$summarize()} method.

<<ictstm_summarize>>=
ce_sim_ictstm <- ictstm$summarize()
head(ce_sim_ictstm$costs)
@

The output is a generic cost-effectiveness object (of class \code{ce}) that can be used to perform CEA. We demonstrate a formal CEA in Section~\ref{sec:cea}. Here, we make a simple pairwise comparisons between each of the new interventions (New 2 and New 1) and the comparator (SOC), and compute the incremental cost-effectiveness ratio (ICER).
